\textcolor{red}{Simple} connectome inference using partial correlation from imaging calcium signal

### 2014-06-16 article.tex
introduction wording to define some notions

% Introduce notation + goal in mathematical score
% TODO check convention for y_{ij}
The connectomes, the neural network, is represented by a (un)directed graph $G = (V, E)$ comprising
a set of nodes $V$ and a set of edges $E \subseteq \left\{(i, j) \in V \times
V\right\}$, (un)ordered pair from $V$.
Each node $i \in V$ represents a neuron, and each edge $(i, j) \in E$
represents a direct link and interactions between neuron $i$ and a neuron $j$.
In the directed graph, each edges also expresses a causal relationship: an
edge $(i, j) \in E$ indicates that the state of neuron $j$ might be caused
by the neuron $i$. The network inference task is set as follows :
\textit{Given the observations $(x^t \in \mathcal{R}^{|V|})_{t=1}^T $
along $T$ periods of time of the $|V|$ neurons, the goal is to infer the
adjacency matrix $A : a_{i,j} = \mathbb{1}((i, j) \in E)$ of the neural network
of graph $G$, where $\mathbb{1}$ is the indicator function.}

@@

\textcolor{red}{The raw signal is the calcium fluorescence imaging representing
neuron activity. Considering the imaging process, we have to cope with a slower
sampling rate than the neuron firing frequency and the slow decay of fluorescent
calcium indicator after a neuron activity burst.}

\begin{wrapfigure}{r}{0.7\linewidth}
\caption{Filtering steps.}\label{wrap-fig:1}
\includegraphics[width=\linewidth]{images/fig_filtering_recut.png}
\end{wrapfigure}

@@

Prior inferring the network, the fluorescent calcium time series
are smoothed using several linear and non linear filters \textcolor{red}{in
order to reduce the imaging process effects. We successively apply different
filters listed below.} \\

@@

\textcolor{red}{As it can be seen on \textit{fig 1}, the raw signal is very noisy and
a low-pass filter is therefore applied to get a cleaner version of the signal where only
true variations are kept, i.e. without noise variations. However, the choice of a low-pass
filter is not unique. In the rest of the paper, we will focus on two types, the
symmetrical median filter, i.e., $X_t^i \leftarrow X_{t-1}^i + X_{t}^i + X_{t+1}^i$,
and an asymmetrical weighted median filter, i.e.,
$X_t^i \leftarrow 0.4 X_{t-3}^i + 0.6 X_{t-2}^i + 0.8 X_{t-1}^i + X_{t}^i$.}\\

\textcolor{red}{Considering the short delay for a communication between
neurons, slow events can be neglected to retrieve direct relationships.
Indeed, for a direct relationship, i.e. a neuron firing directly implying another
neuron to fire, the average time is about $1ms$ or $2ms$ while the
fluorescent signal time resolution is $20ms$. In order to find out which neurons
are directly related, it is only necessary to look after the burst signal, or
in other words, whether the signal varies. In our approach, we choose to apply
an asymmetrical discrete derivative : $X_{t,d}^{i} \leftarrow X_{t}^i - X_{t-1}^i$.}\\

\textcolor{red}{By nature, the firing event is really fast and in order to
retrieve the connectivity from them, it is necessary to link a burst to a
couple of timesteps (ideally one). The small decay for the fluorescent calcium
indicator implies a signal variation during a long interval and therefore, in
order to only keep rising and strong transitions, we apply a hard-tresholing filter:
$X_{t,d}^i = X_{t,d}^i \mathbb{1}(X_{t,d}^i \geq \tau)$ where $\tau$ is a given threshold.
As a result, we get a binary-like signal with the true activity when a burst
happens and with a true zero when the neuron is inactive.}

@@

\textcolor{red}{Then we apply on the entire signal a non-linear filter based on the
current activity of the systems in order to favor the association measure inversely
proportional to the number of neurons firing together:
$ X_{t,w}^i  = (X_{t,d}^i + 1 )^{1 + \frac{1}{\sum_{j} X_{t,d}^j}} $.
When too much neurons are firing together, i.e. in case of a network burst, it is almost
impossible to retrieve direct relationships: all neurons seem to be directly
connected to each other. By favoring the association measure when the activity is low,
we expect to retrieve a pair of neurons that are really directly connected together.
In concrete terms, as it can be verified on \textit{fig 1}, peak intensity may change in function
of the activity of all neurons but the number and the locations of peaks is not modified.
}

>>>>

% Introduce notation + goal in mathematical score
% TODO check convention for y_{ij}
The connectomes, the neural network, is represented by a (un)directed graph $G = (V, E)$ comprising
a set of nodes $V$ and a set of edges $E \subseteq \left\{(i, j) \in V \times
V\right\}$, (un)ordered pair from $V$.
Each node $i \in V$ represents a neuron, and each edge $(i, j) \in E$
represents a direct link from a neuron $i$ to a neuron $j$.
The network inference task is set as follows :
\textit{Given the observations $(x^t \in \mathcal{R}^{|V|})_{t=1}^T $
along $T$ periods of time of the $|V|$ neurons, the goal is to infer the
adjacency matrix $A : a_{i,j} = \mathbb{1}((i, j) \in E)$ of the neural network
of graph $G$, where $\mathbb{1}$ is the indicator function.}

The directivity in the network introduce causal relations through neurons. A directed edge, i.e. $a_{i,j} = 1$ and $a_{j,i} = 0$, means that neuron $i$ causes $j$. The causal mechanism implies that when neuron $i$ is active, also denoted in this paper as a firing event or a burst, neuron $j$ will be consequently active aftewards. The reciprocal phenomenon is only possible when the edge is bidirectionnal, i.e. $a_{i,j} = 1 = a_{j,i} = 1$.

@@

The raw signal is the calcium fluorescence imaging representing
neuron activity. Considering the imaging process, we have to cope with a slower
sampling rate than the neuron firing frequency and the slow decay of fluorescent
calcium indicator after a neuron activity burst.

\begin{figure}%{r}{0.7\linewidth}
\caption{Filtering steps.}\label{wrap-fig:1}
\includegraphics[width=\linewidth]{images/fig_filtering_recut.png}
\end{figure}

@@

Prior inferring the network, the fluorescent calcium time series
are smoothed using several linear and non linear filters in
order to reduce the imaging process effects. We successively apply different
filters listed below.

@@

As it can be seen on \textit{fig 1}, the raw signal is very noisy and
a low-pass filter is therefore applied to get a cleaner version of the signal where only
true variations are kept, i.e. without noise variations. However, the choice of a low-pass
filter is not unique. In the rest of the paper, we will focus on two types, the
symmetrical median filter, i.e., $x^t_i \leftarrow x^{t-1}_i + x^{t}_i + x^{t+1}_i$,
and an asymmetrical weighted median filter, i.e.,
$x^t_i \leftarrow 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_{t}^i$.

Considering the short delay for a communication between
neurons, slow events can be neglected to retrieve direct relationships.
Indeed, for a direct relationship, i.e. a neuron firing directly implying another
neuron to fire, the average time is about $1ms$ or $2ms$ while the
fluorescent signal time resolution is $20ms$. In order to find out which neurons
are directly related, it is only necessary to look after the burst signal, or
in other words, whether the signal varies. In our approach, we choose to apply
an asymmetrical discrete derivative : $x^{t,d}_{i} \leftarrow x^{t}_i - x_{t-1}^i$.

By nature, the firing event is really fast and in order to
retrieve the connectivity from them, it is necessary to link a burst to a
couple of timesteps (ideally one).

The small decay for the fluorescent calcium
indicator implies a signal variation during a long interval and therefore, in
order to only keep rising and strong transitions, we apply a hard-tresholing filter:
$x^{t,d}_i = X^{t,d}_i \mathbb{1}(x^{t,d}_i \geq \tau)$ where $\tau$ is a given threshold 
As a result, we get a binary-like signal with the true activity when a burst
happens and with a true zero when the neuron is inactive.

@@

Then we apply on the entire signal a non-linear filter based on the
current activity of the systems in order to favor the association measure inversely
proportional to the number of neurons firing together:
$ X_{t,w}^i  = (x^{t,d}_i + 1 )^{1 + \frac{1}{\sum_{j} X_{t,d}^j}} $.
When too much neurons are firing together, i.e. in case of a network burst, it is almost
impossible to retrieve direct relationships: all neurons seem to be directly
connected to each other. By favoring the association measure when the activity is low,
we expect to retrieve a pair of neurons that are really directly connected together.
In concrete terms, as it can be verified on \textit{fig 1}, peak intensity may change in function
of the activity of all neurons but the number and the locations of peaks is not modified.
